Generative Adversarial Networks

### 2017-12-03 cs6735/project/report.tex
more edits

\subsection{Related Work}
Until the advent of GANs, most deep generative models provided a parametric specification of an probability distribution function \citep{2014arXiv1406.2661G}. These models could be trained using maximum likelihood estimation which often required gradient estimation. The most successful of these models is the deep Boltzmann machine \citep{2014arXiv1406.2661G}. The difficulties associated with gradient estimation motivated the development of \texit{generative machines}, models which implicitly represent the joint probability distribution by generating samples from that distribution rather than explicitly rendering a probability distribution.

@@

\subsection{Instability}
Generative adversarial networks are often characterized by their instability \citep{2017arXiv170100160G}. The training process makes use of gradient descent which does not guarantee convergence to the Nash equilibrium, the game theory equilibrium that involves two competing players. This training process for GANs is different than network optimization which is common in discriminative models. When either the generator or discriminator trains, even if that specific network improves, the other may regress. This sometimes leads to the scenario where each network repeatedly undo's the other's progress. This problem is not specific to GANs, but to any zero-sum game situations \citep{2017arXiv170100160G}.

@@

\subsection{Image Quality}
When generating images, problems often arise due to the generators difficulties with perspective, counting and structure \citep{2017arXiv170100160G}. For example, an generated animal may not have a 3D structure, have misplaced limbs or have too many of a specific body part. These issues arise due to the underlying model architecture and are not an inherent problem the adversarial network. Furthermore, generated image resolution has remained relatively low; however, recent improvements outlined in the following section show improvements with regard to resolution \citep{2017arXiv171010196K}.

\subsection{Evaluation}
The evaluation procedure for generative adversarial networks remains an open debate as there exist no clear quantitative measurement to evaluate the models \citep{2017arXiv170100160G}. For example, models with high likelihood may produce poor samples and models with low likelihood may produce realistic samples. Several variations of the GANs examine the tradeoffs of divergence measures other than the original Kullback-Leibler divergence \citep{2015arXiv151101844T}. It can be shown that different metrics favor different models. Therefore, it is important to asses one's model within the context of the whole application \citep{2015arXiv151101844T}.

\subsection{Discrete Data}
GANs assume a model implementation that is differentiable due to it's use of back propagation. Therefore, this restricts their ability to generate discrete data. Goodfellow proposes several solutions such as algorithm modifications, discretization and the use of concrete distributions or Gumbel-softmax \citep{2017arXiv170100160G}.

\section{Training Improvements}
A substantial amount of research has been conducted within the past three years with regards to improving GAN training procedures. A paper released by Goodfellow approximately one year after the debut of GANs outlines several techniques that can be applied to improve the training process \citep{2016arXiv160603498S}. As previously mentioned, the optimal solution for GANs remains convergence of the Nash equilibrium of the minimax game; however, gradient descent attempts to find a local minima. The following techniques introduced by Goodfellow \textit{et al.} address the instability of GANs and attempt to increase the probability of convergence.

\subsection{Feature Matching}
Feature matching involves the definition of a new cost function for GANs such that they do not overtrain on the current discriminator \citep{2016arXiv160603498S}. Instead of optimizing on the output of the discriminator, the new objective requires the generator to minimize the difference between the expected value on some intermediate layer. The objective function can be described as $\Big|\Big|\mathbb{E}_{x\sim p_{data}(x)}\Big[f(x)\Big] - \mathbb{E}_{z\sim p_{z}(z)}\Big[f(G(z))\Big]\Big|\Big|_2^2$ where $f(x)$ is the output of the activation function of some layer. Experimental evidence has been gathered that demonstrates the usefulness of this technique, especially when the training process becomes instable \citep{2016arXiv160603498S}. Furthermore, feature matching is useful for semi-supervised GAN applications as will later be described in section \ref{sec:semisupervised}.

\subsection{Minibatch Discrimination} \label{sec:minidisc}
Minibatch discrimination attempts to minimize the issues caused by mode collapse. This problem occurs as the discriminator processes each value $x$ independently of each other value. The gradient of each will point in the same direction thus leading to mode collapse. This technique involves the use of a minibatch in place of a single value to allow to discriminator to examine multiple data examples at one time. The discriminator is then able to examine the distance between each sample in conjunction with the samples themselves and look for abnormal minibatch distributions. Successful application of this technique allows for visually appealing images to be generated much quicker than by previous training techniques \citep{2016arXiv160603498S}. Preliminary results also show great reduction of mode collapse issues \citep{2017arXiv170100160G}.

\subsection{Historical Averaging}
Historical averaging introduces an extra term for each player's cost function such that they are penalized when choosing new parameters that differ from the historical mean value \citep{2016arXiv160603498S}. This new term $||\theta - \frac{1}{t}\sum_{i=1}^{t}\theta[i]||^2$ where $t$ is the total number of iterations performed and $\theta[i]$ is the value of the parameters at iteration $i$. This technique has been demonstrated to help low dimensionality games reach an equilibrium where normal gradient descent has failed \citep{2016arXiv160603498S}.

\subsection{Label Smoothing}
Label smoothing was a technique first introduced in the 1980s which involves replacing the 0 and 1 targets with \textit{smoother} values such as 0.1 and 0.9 respectively \citep{2016arXiv160603498S}. This technique is used to improve the networks resistance to adversarial examples. Specifically, it is recommendended to use one-sided label smoothing by using 0.9 label instead of 1 for the real data. By altering the label for fake data, this may lead to problems when $p_{data}(x)$ nears $0$ and $p_{model}(x)$ becomes large for some sample $x$ \citep{2016arXiv160603498S}.

\subsection{Virtual Batch Normalization}
Batch normalization is a technique which involves applying normalization for each training minibatch. This allows for much higher learning rates, more stable training and reduces problems associated with parameter initialization \citep{2015arXiv150203167I}. This normalization technique was shown to successful for Deep Convolutional GANs; however, it causes a dependence relationship between the samples of the minibatch \citep{2016arXiv160603498S}. When implementation virtual batch normalization, a reference batch of samples is randomly chosen at the start of training and used throughout to normalize the current sample. However, this technique is computationally expensive as it requires forwarding two minibatches.

\section{Architecture Variations}
Research within the field of deep generative modeling has increased exponentially since the debut of generative adversarial networks in 2014. Significant improvements are being made with respect to image quality and training stability. Hundreds of papers have been released since 2014 as seen in figure \ref{fig:CumulativeNumberofGANPapersbyMonth} which propose numerous architecture variations.

@@

\subsection{Fully Connected GANs}
The first generative adversarial network paper by Goodfellow \textit{et al.} made use of relatively simple fully connected neural networks for the generator and discriminator models. These are networks where each neuron in one layer is connected to every neuron in the following layer. This design is able to generate simple images and useful for proof of concept; however, several other architectures have overtaken this implementation \citep{2017arXiv171007035C}.

@@

\subsection{Progressive GANs}
Progressive GANs offer extend the DCGAN architecture by altering their training process. Introduced in October 2017 by Karras \textit{et al.} \citep{2017arXiv171010196K}, these networks incrementally increase the number of convolutions within both the generator and discriminator. Each new convolution is \texit{blended in} and is added after the current GAN converges which permits the generator to progressively discover the image distribution. Combined with new normalization techniques, progressive GANs allow for much higher resolution images to be generated \citep{2017arXiv171010196K}. The experimental results offer promising new directions for GAN research and applications.

\begin{figure}
\centering
\includegraphics[scale=0.45]{progressive_growing}
\caption{The training process of progressively growing generative adversarial networks recently introduced by Karras \texit{et al.} \citep{2017arXiv171010196K}. The generator and discriminator begin at $4x4$ resolution and progressively grow generate and discriminate $1024x1024$ resolution images.}
\label{fig:ProgressiveGrowingGAN}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.45]{smile_concept_vector}
\caption{The addition of the "smile" concept vector (far right) to a vector associated with a woman (far left). The images in-between were created through an interpolation of the two vectors. Example retrieved form Dumoulin \texit{et al.} \citep{2016arXiv160600704D}.}
\label{fig:conceptvector}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.7]{multimodal}
\caption{An example of scenario where the ability to model multi-modal data is important. This image demonstrates the process of predicting the next video frame given the previous one. A generator trained with mean squared error (middle) produces an image which averages each possible future. Using an adversarial process (right), the image produced corresponds with the most likely scenario. Example retrieved form Lotter \texit{et al.} \citep{2015arXiv151106380L}.}
\label{fig:multimodal}
\end{figure}

>>>>

\subsection{Related Work}
Until the advent of GANs, most deep generative models provided a parametric specification of an probability distribution function \citep{2014arXiv1406.2661G}. These models could be trained using maximum likelihood estimation which often required gradient estimation. The most successful of these models is the deep Boltzmann machine \citep{2014arXiv1406.2661G}. The difficulties associated with gradient estimation motivated the development of \textit{generative machines}, models which implicitly represent the joint probability distribution by generating samples from that distribution rather than explicitly rendering a probability distribution.

@@

\subsection{Instability} \label{sec:instability}
Generative adversarial networks are often characterized by their instability \citep{2017arXiv170100160G}. The training process makes use of gradient descent which does not guarantee convergence to the Nash equilibrium, the game theory equilibrium that involves two competing players. This training process for GANs is different than network optimization which is common in discriminative models. When either the generator or discriminator trains, even if that specific network improves, the other may regress. This sometimes leads to the scenario where each network repeatedly undo's the other's progress. This problem is not specific to GANs, but to any zero-sum game situations \citep{2017arXiv170100160G}.

@@

\subsection{Image Quality}
When generating images, problems often arise due to the generators difficulties with perspective, counting and structure \citep{2017arXiv170100160G}. For example, an generated animal may  have a 2D structure, misplaced limbs or too many of a specific body part. These issues arise due to the underlying model architecture, typically convolutional neural networks, and are not a direct problem of the adversarial network \citep{2017arXiv170100160G}. Furthermore, generated image resolution has remained relatively low; however, recent improvements outlined in section \ref{sec:proggans} show improvements with regard to resolution.

\subsection{Evaluation}
The evaluation procedure for generative adversarial networks remains an open debate as there exist no clear quantitative measurement to evaluate the models \citep{2017arXiv170100160G}. For example, models with high likelihood may produce unrealistic samples and models with low likelihood may produce realistic samples \citep{2017arXiv170100160G}. Several variations of the GANs examine the tradeoffs of divergence measures other than the original Kullback-Leibler divergence \citep{2015arXiv151101844T}. It has been shown that different metrics lead to different tradeoffs and different evaluations favor different models \citep{2015arXiv151101844T}. Therefore, it is important to asses one's model within the context of the whole application \citep{2015arXiv151101844T}.

\subsection{Discrete Data}
GANs assume a model implementation that is differentiable due to it's use of back propagation. Therefore, this restricts their ability to generate discrete data. Goodfellow proposes several solutions such as algorithm modifications, discretization of outputs or the use of concrete distributions or Gumbel-softmax \citep{2017arXiv170100160G}.

\section{Training Improvements}
A substantial amount of research has been conducted within the past three years with regards to improving GAN training procedures. A paper released by Salimans \textit{et al.} while at OpenAI outlines several techniques that can be applied to improve the GAN training process \citep{2016arXiv160603498S}. As mentioned in section \ref{sec:instability}, the optimal solution for GANs is the convergence to the Nash equilibrium of the minimax game; however, gradient descent only attempts to minimize the cost function and not to find the Nash equilibrium. The following techniques introduced in the paper address the instability of GANs and attempt to increase the probability of convergence.

\subsection{Feature Matching}
Feature matching involves the definition of a new cost function for GANs so that the generator does not overtrain on the current discriminator \citep{2016arXiv160603498S}. Instead of optimizing on the output of the discriminator, the new objective requires the generator to minimize the difference between the expected value of some intermediate layer. The objective function can be described as $\Big|\Big|\mathbb{E}_{x\sim p_{data}(x)}\Big[f(x)\Big] - \mathbb{E}_{z\sim p_{z}(z)}\Big[f(G(z))\Big]\Big|\Big|_2^2$ where $f(x)$ is the output of the activation function of some layer of the discriminator. Experimental evidence gathered has demonstrated the usefulness of this technique, especially when the training process becomes instable \citep{2016arXiv160603498S}. Furthermore, feature matching is useful for semi-supervised GAN applications as will later be described in section \ref{sec:semisupervised}.

\subsection{Minibatch Discrimination} \label{sec:minidisc}
Minibatch discrimination attempts to minimize the issues caused by mode collapse. This problem occurs as the discriminator processes each value $x$ independently of each other value. The gradients of discriminator will all point in the same direction, thus leading to mode collapse. This technique involves the use of a minibatch in place of a single value to allow to discriminator to examine multiple data examples at one time. The discriminator is then able to examine the distance between each sample in conjunction with the samples themselves and look for abnormal minibatch distributions. Successful applications of this technique allow for visually appealing images to be generated much quicker than by previous training techniques \citep{2016arXiv160603498S}. Preliminary results also show great reduction of mode collapse issues \citep{2017arXiv170100160G}.

\subsection{Historical Averaging}
Historical averaging introduces an extra term for each player's cost function such that they are penalized when choosing new parameters that differ from the historical mean value \citep{2016arXiv160603498S}. The new term becomes $||\theta - \frac{1}{t}\sum_{i=1}^{t}\theta[i]||^2$ where $t$ is the total number of iterations performed and $\theta[i]$ is the value of the parameters at iteration $i$. This technique has been demonstrated to help low dimensionality games reach an equilibrium where normal gradient descent has failed \citep{2016arXiv160603498S}.

\subsection{Label Smoothing}
Label smoothing was a technique first introduced in the 1980s which involves replacing the 0 and 1 targets with \textit{smoother} values such as 0.1 and 0.9 respectively \citep{2016arXiv160603498S}. This technique is used to improve the networks resistance to adversarial examples. Specifically, it is recommended one use only one-sided label smoothing by using 0.9 label instead of 1 for the real data label. If one alters the label for fake data, problems may arise when $p_{data}(x)$ nears $0$ and $p_{model}(x)$ becomes large for some sample $x$ \citep{2016arXiv160603498S}.

\subsection{Virtual Batch Normalization}
Batch normalization is a technique that was first introduced in 2014 which involves applying normalization for each training minibatch \citep{2015arXiv150203167I}. It allows for much higher learning rates, a more stable training process and a reduction of problems associated with parameter initialization \citep{2015arXiv150203167I}. This normalization technique has proved useful for GAN training; however, it often causes a dependence relationship between the samples of the minibatch \citep{2016arXiv160603498S}. A new technique called virtual batch normalization (VBN) was developed where a reference batch of samples is randomly chosen at the start of the training process and used throughout to normalize each training sample. However, this technique is computationally expensive as it requires forwarding two minibatches and it is recommended to only use VBN for the generator \citep{2016arXiv160603498S}.

\section{Architecture Variations}
Research within the field of deep generative modeling has increased exponentially since the debut of generative adversarial networks in 2014. Significant improvements are being made with respect to image quality and training stability. Hundreds of papers have been released since 2014 (see figure \ref{fig:CumulativeNumberofGANPapersbyMonth}) which propose numerous architectural variations.

@@

\subsection{Fully Connected GANs}
The first generative adversarial network paper by Goodfellow \textit{et al.} made use of relatively simple fully connected neural networks for the generator and discriminator models \citep{2017arXiv171007035C}. These are networks where each neuron in one layer is connected to every neuron in the following layer. This design is able to generate simple images and useful for proof of concept; however, several other architectures have overtaken this implementation \citep{2017arXiv171007035C}.

@@

\subsection{Progressive GANs} \label{sec:proggans}
Progressive GANs offer extend the DCGAN architecture by altering their training process. Introduced in October 2017 by Karras \textit{et al.} \citep{2017arXiv171010196K}, these networks incrementally increase the number of convolutions within both the generator and discriminator. Each new convolution is \textit{blended in} and is added after the current GAN converges which permits the generator to progressively discover the image distribution. Combined with new normalization techniques, progressive GANs allow for much higher resolution images to be generated \citep{2017arXiv171010196K}. The experimental results offer promising new directions for GAN research and applications.

\begin{figure}
\centering
\includegraphics[scale=0.45]{progressive_growing}
\caption{The training process of progressively growing generative adversarial networks recently introduced by Karras \textit{et al.} \citep{2017arXiv171010196K}. The generator and discriminator begin at $4x4$ resolution and progressively grow generate and discriminate $1024x1024$ resolution images.}
\label{fig:ProgressiveGrowingGAN}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.40]{smile_concept_vector}
\caption{The addition of the "smile" concept vector (far right) to a vector associated with a woman (far left). The images in-between were created through an interpolation of the two vectors. Example retrieved form Dumoulin \textit{et al.} \citep{2016arXiv160600704D}.}
\label{fig:conceptvector}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.7]{multimodal}
\caption{An example of scenario where the ability to model multi-modal data is important. This image demonstrates the process of predicting the next video frame given the previous one. A generator trained with mean squared error (middle) produces an image which averages each possible future. Using an adversarial process (right), the image produced corresponds with the most likely scenario. Example retrieved form Lotter \textit{et al.} \citep{2015arXiv151106380L}.}
\label{fig:multimodal}
\end{figure}
